### 2018-11-13 article/ms.tex
words

\documentclass[twocolumn]{aastex61}
\usepackage{bm}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\usepackage{comment}

\newcommand\teff{T_{\rm eff}}
\newcommand\logg{\log{g}}
\newcommand\feh{[\rm{Fe}/\rm{H}]}

\newcommand{\project}[1]{\textsl{#1}}
\newcommand{\package}[1]{\texttt{#1}}
\newcommand{\acronym}[1]{{\small{#1}}}

\newcommand{\Gaia}{\project{Gaia}}
\newcommand{\gaia}{\project{gaia}}
\newcommand{\Galah}{\project{Galah}}
\newcommand{\GALAH}{\project{GALAH}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\newcommand{\vect}[1]{\boldsymbol{\mathbf{#1}}}
\renewcommand{\vec}[1]{\vect{#1}}

\newcommand{\weight}{\pi}
\newcommand{\data}{\textbf{Y}}
\newcommand{\vecdata}{\vec\data}

\newcommand{\nextstep}{^\textrm{(t+1)}}
\newcommand{\thisstep}{^\textrm{(t)}}
\newcommand{\transpose}{^\intercal}
\newcommand{\eye}{\textbf{I}}

\newcommand{\factorloads}{\textbf{L}}
\newcommand{\factorscores}{\textbf{S}}
\newcommand{\specificvariance}{\vec{D}}

\newcommand{\scoremeans}{\vec\xi}
\newcommand{\scorecovs}{\vec\Omega}

\received{2018 XX XX}
\revised{2018 XX XX}
\accepted{2018 XX XX}

\newcommand{\vcpath}{vc.tex}

\IfFileExists{\vcpath}{\input{\vcpath}}{
	\newcommand{\giturl}{UNKNOWN}
	\newcommand{\gitslug}{UNKNOWN}
	\newcommand{\githash}{UNKNOWN}
	\newcommand{\gitdate}{UNKNOWN}
	\newcommand{\gitauthor}{UNKNOWN}
}

\submitjournal{AAS Journals}

\shorttitle{Statistically consistent chemical tagging}
\shortauthors{Casey et al.}

\begin{document}

\title{A statistically consistent approach to chemical tagging}

\correspondingauthor{Andrew R. Casey}
\email{andrew.casey@monash.edu}

\author[0000-0003-0174-0564]{Andrew R. Casey}
\affiliation{School of Physics \& Astronomy, 
			 Monash University,
			 Wellington Rd, Clayton 3800, Victoria, Australia}
\affiliation{Faculty of Information Technology, 
			 Monash University, 
			 Wellington Rd, Clayton 3800, Victoria, Australia}

\author{John Lattanzio}
\affiliation{School of Physics \& Astronomy, 
			 Monash University,
			 Wellington Rd, Clayton 3800, Victoria, Australia}

\author{Aldeida Aleti}
\affiliation{Faculty of Information Technology, 
			 Monash University, 
			 Wellington Rd, Clayton 3800, Victoria, Australia}

\author{David Dowe}
\affiliation{Faculty of Information Technology, 
			 Monash University, 
			 Wellington Rd, Clayton 3800, Victoria, Australia}

\author{the GALAH team}

\begin{abstract}
\end{abstract}

\keywords{methods: statistical}

\section{Introduction} \label{sec:intro}

\begin{itemize}
	\item Chemical tagging
	\item Limits on the efficacy of chemical tagging
	\item Clustering vs common traits
\end{itemize}

The \Galah\ survey \citep{DaSilva:2015} is a stellar spectroscopic survey 
that uses the High Efficiency and Resolution Multi-Element 
Spectrograph \citep[HERMES][]{Sheinis:2016} on the Australian Astronomical Telescope (AAT).
The survey will observe up to $10^6$ stars in the Milky Way, and
measure up to 30 chemical abundances for each star. This includes
light odd-Z elements (e.g., Na, K), elements produced through
alpha-particle capture (e.g., Mg, Ca, Ti), and elements produced
through the slow (e.g., Ba) and rapid neutron-capture process
(e.g., Eu). These data provide an unparalleled view on the production
of chemical elements in the Milky Way.

\section{Methods} \label{sec:methods}

% TODO: Is D a diagonal vector, or a matrix?! Make it consistent.

% Describe how we will attempt to define a generative model for the data, from the latent space.

In this work the data $\vecdata$ is a $n \times d$ matrix 
where $n$ is the number of stars and $d$ is the number of
chemical abundances measured for each star. We assume a
generative model for the data 
\begin{equation}
	\vecdata = \factorloads\factorscores + \vec{e}
	\label{eq:generative-model}
\end{equation}

\noindent{}where $\factorloads$ is a $p \times d$ matrix of factor
loads that is common to all data points, and the factor scores for
the $i$th data point
\begin{equation}
	\factorscores_i \sim \mathcal{N}(\vec\xi_k, \vec\Omega_k)
\end{equation}
\noindent{}are drawn from the $k$th multivariate  normal distribution.
The factor scores for all data points $\factorscores$ is then a 
\todo{$n \times p$} matrix.
 We assume $\vec{e} \sim \mathcal{N}\left(\vec{0}, \eye\specificvariance\right)$
is independent of the latent space, and $\specificvariance$ is a
diagonal matrix of $d$ entries.

In this model each data point can be represented as being drawn
from a mixture of multivariate normal components, except the components
are \emph{clustered in the latent space} $\factorscores$ and projected
into the data space by the factor loads $\factorloads$. 
We assume that the latent space is lower dimensionality than the
data space (e.g., $p < d$).
Within the context of stellar abundances, the factor loads
$\factorloads$ can be thought of as the \emph{mean} yields
of nucleosynthetic
events (e.g., $s$-process yields from AGB stars), and the
factor scores are analogous to the relative counts of those 
nucleosynthetic events. The clustering in factor scores
achieves the same as a clustering procedure in data space,
except we simultaneously estimate the factor loads
(nucleosynthetic yields) that are common to all data points 
(stars). Within this framework a rare nucleosynthetic event
can still be described as a `common factor load' $\factorloads_i$, 
but its rarity would be represented by associated factor
scores being zero for most stars and thus have no contribution
to the observed abundances. The factor loads and factor scores 
cannot be expressly interpreted like this because they have no
physical meaning, but this description of relative yields and 
rates should help build intuition for the model parameters,
and provide an astrophysical context.

One benefit of using a mixture of common factor analysers is
for modelling stellar chemical abundances is the scaling with
computational cost. If we considered data sets of order $3\times10^7$
entries (e.g., 30 chemical abundances for $10^6$ stars) purely as a
clustering problem, then even the most efficient clustering
algorithms would incur a significant cumulative computational 
overhead by searching the parameter space for the number of
clusters, and the optimal model parameters given that number
of components. However, because the mixture of factor analyzers
approach assumes that there is a \emph{lower dimensional latent 
space} in which the data are clustered, and that clustering is 
projected into real space by common factor loads, the 
dimensionality of the clustering problem is reduced from 
$N \times D$ to $N \times J$. This reduces cost both through
faster execution of each E-M step, and on average fewer E-M steps
required to reach a specified convergence threshold.

From a statistical standpoint, the primary advantage to using
a mixture of factor analysers is that we can simultaneously
estimate latent factors (e.g., infer nucleosynthetic 
yields) and perform clustering (e.g., chemical tagging) 
within a statistically consistent framework. That is to say
that we have a generative model for the data that includes
analogues of nucleosynthetic yields and star formation,
and the parameters of this model can be estimated consistently
given some data.

Without loss of generality the density of the data $\vecdata$ can be modelled as
\begin{equation}
	f(\textbf{y}; \vec\Psi) = \sum_{k=1}^{K}\weight_{k}\phi(\textbf{y};\factorloads\scoremeans_k, \factorloads\scorecovs_k\factorloads\transpose + \eye\specificvariance)
\end{equation}
\noindent{}given $p$ common factor loadings and $K$ components
clustered in the latent (factor score) space. Here the parameter
vector
$\vec\Psi$ includes $\{\factorloads,\vec\pi,\scoremeans,\scorecovs,\specificvariance\}$, and $\phi(\textbf{y};\vec\mu, \vec\Sigma)$
describes the density of a multivariate gaussian distribution with
mean $\vec\mu$ and covariance matrix $\vec\Sigma$,
and $\weight_k$ describes the relative weighting of the $k$th
component in latent space and $\sum\weight_k = 1$.
The log likelihood is then given by
\begin{equation}
	\log\mathcal{L}(\vecdata|\vec\Psi) = \sum_{k=1}^{K}\log{f(\vecdata;\vec\Psi)} \quad .
\end{equation}

The model described by Equation~\ref{eq:generative-model} is indeterminate:
there is no unique solution for the factor loads $\factorloads$ and scores
$\factorscores$. However, an accurate estimate of the parameter vector $\vec\Psi$ can be obtained by the expectation-maximization algorithm \citep{EM}.

\subsection{Initialisation}

We initialise the entries in the factor loads $\factorloads$ by drawing
from a uniform distribution $\factorloads \sim \mathcal{U}\left(-1, +1\right)$.
We assign each data point as belonging to one of the $K$ components
by drawing from a multinomial distribution. Given these assignments and
factor loads, we then estimate the relative weights $\vec\pi$, the 
mean factor scores of each component $\scoremeans$, and the covariance
matrix of factor scores of each component $\scorecovs$.

\begin{comment}
We use the K-means++ algorithm to initially assign data points
randomly to $K$ clusters (in data space). This provides a so-called
hard initial association where each data point is wholly assigned
to a single component. Specifically we calculate the $n \times K$
responsibility matrix $\vec\tau$, where a entry of 1 indicates the
$n$th data point is associated with the $k$th cluster, and an
entry of 0 indicates no association such that
\begin{equation}
	\weight_k = \frac{1}{N}\sum_{n=1}^{N}\tau_{nk} \quad .
\end{equation}
We then chose 1,000
random data points and factorize that matrix using singular-value
decomposition. We take the principal $p$ eigenvectors as the 
initial guess for the factor loads $\factorloads$, and calculate
an initial estimate of the factor score means
\begin{equation}
	\scoremeans_k = \frac{1}{N}\sum_{n=1}^{N}\tau_{nk}\vecdata_{n}\factorloads
\end{equation}
\noindent{}and the covariance matrices of the clusters in latent space
\begin{equation}
	\scorecovs_k = \textrm{cov}{\left(\tau_{nk}\vecdata_{n}\right)} \quad .
\end{equation}
\end{comment}

\subsection{Expectation-Maximization}

At the expectation step we evaluate the log likelihood
given the model parameters $\vec\Psi$, and we re-calculate the responsibility
matrix whose entries are the posterior probability
that the $n$th data point is associated to the $k$th component,
given the data $\vecdata$ and the current estimate of the 
parameter vector $\vec\Psi$:
\begin{equation}
	\tau_{nk} = \frac{\weight_k\phi(\vecdata_n;\factorloads\scoremeans_k, \factorloads\scorecovs_k\factorloads\transpose + \eye\specificvariance)}{\sum_{g=1}^{G}\weight_g\phi(\vecdata_n;\factorloads\scoremeans_g, \factorloads\scorecovs_g\factorloads\transpose + \eye\specificvariance)} \quad .
\end{equation}

% The complexity of SVD is O(n^3), and
% the complexity of K-means++ is O(log(k)). If we knew the complexity of
% k-means++ relative to N, then we would be able to make better decisions
% as to how to initialise: SVD first and then K-means in latent space?

% Discuss connection to PCA, FA, MFA.

% Include macros for \vec, \data, etc.
At the maximization step we update our estimates of the parameters,
conditioned on the data $y$ and the responsibility matrix $\vec\tau$.
Initially we update the relative weights $\vec\weight\nextstep$ given
the responsibility matrix $\vec\tau$
\begin{equation}
	\weight_k\nextstep = \frac{1}{N} \sum_{n=1}^{N}\tau_{nk} \quad .
	% Include proof.
\end{equation}

%\noindent{}In the equations that follow (Equation~\ref{eq:X} to \ref{eq:Y})
%when we refer to any value in $\{\factorloads,\vec\pi,\scoremeans,\scorecovs,\specificvariance\}$ for brevity we are referring to the current estimate of that
%parameter (e.g., $\factorloads\thisstep$). The updated estimates are marked as such
%(e.g., $\scoremeans\nextstep$).

The updated estimates of the mean factor scores 
$\scoremeans\nextstep$ are given by
\begin{eqnarray}
	\scoremeans_{k}\nextstep = \scoremeans_{k}\thisstep + \frac{\vec{G}\transpose(\vecdata\transpose - \factorloads\thisstep\scoremeans_{k}\thisstep)\vec\tau_{k}}{N\weight_k\nextstep}
\end{eqnarray}
\noindent{}where:
\begin{eqnarray}
	\vec{W} &=& (\scorecovs_{k}\thisstep)^{-1}\eye \\
	\vec{V} &=& \left(\specificvariance\thisstep\right)^{-1} \\
	\vec{C} &=& (\vec{W} + (\factorloads\thisstep)\transpose\vec{V}\factorloads\thisstep)^{-1}\eye \\
	\vec{G} &=& \left[\vec{V} - \vec{V}\factorloads\thisstep\vec{C}\left(\vec{V}\factorloads\thisstep\right)\transpose\right]\factorloads\thisstep\scorecovs_k\thisstep \quad .
\end{eqnarray}

The covariance matrices of the components of factor scores $\scorecovs\nextstep$
are updated next,
\begin{equation}
	\scorecovs_k\nextstep = \left(\eye - \vec{G}\transpose\factorloads\thisstep\right)\scorecovs_k\thisstep + \frac{\vec{G}\transpose\vec{Z}\left(\vec{Z}\vec\tau_k\transpose\right)\transpose\vec{G}}{N\weight_k\nextstep}
\end{equation}
\noindent{}where
\begin{eqnarray}
	\vec{Z} &=& \vecdata\transpose - \factorloads\thisstep\scoremeans_k\thisstep \quad .
\end{eqnarray}

After some linear algebra, updated estimates of the common factor loads $\factorloads\thisstep$
can be found from
\begin{equation}
	\factorloads\nextstep = \factorloads_{1}\left(\factorloads_{2}^{-1}\eye\right)
\end{equation}
\noindent{}where:
\begin{eqnarray}
	\factorloads_1 &=& \sum_{k=1}^{K}\left[ \vec\tau_k\transpose\vecdata\left(\scoremeans_k\thisstep\right)\transpose + 
	\vec{G}\transpose\vec\tau_k\vec{Z}\transpose\vec{G}\right] \\
	%\left(\scoremeans_k\thisstep\vecdata\transpose\vec\tau_k\right)\transpose \left(\vec{G}\transpose\vec{Z}\vec\tau_k\transpose\vec{G}\right)\transpose\right]  \\
	\factorloads_2 &=& N\sum_{k=1}^{K}\left[\weight_k\nextstep\left(\scorecovs_k\nextstep + \scoremeans_k\nextstep\left(\scoremeans_k\nextstep\right)\transpose\right)\right]
\end{eqnarray}

Finally, the updated estimate of the specific variances $\specificvariance\nextstep$ is given
by
\begin{equation}
	\specificvariance\nextstep = \frac{1}{N}\left[\sum^{K}_{k=1}\vec\tau_k\transpose\left(\vecdata\odot\vecdata\right) - \sum_{j=1}^{J}\left(\factorloads\nextstep\factorloads_2\right)\odot\factorloads\nextstep\right]
\end{equation}

\noindent{}where $\odot$ denotes is the entrywise (Hadamard) product.

\section{Experiments} \label{sec:experiments}

\subsection{Experiment 1: Toy model with generated data} \label{sec:experiment-toy-model}

We generated a data set with ${N = 100000}$ data points, each with
$D = 15$ dimensions, and assumed that those data are generated by
latent space of $J = 5$ ($J \times D$) factor loads, and there are $K = 20$
clusters in the latent space. The relative weights $\vec\weight$
are drawn from a multinomial distribution and the means of the clusters
in factor scores $\scoremeans$ are drawn from a standard normal
distribution. The covariance matrices in factor scores $\scorecovs$ are assumed to be diagonal matrices, where the non-zero entries are drawn from a gamma distribution $\scorecovs_{k,i,i} \sim \vec\Gamma\left(1\right)$. The variance in 
each dimension $\specificvariance$ are also drawn $\specificvariance \sim \vec\Gamma\left(1\right)$.
The $i$th data point (which belongs to the $k$th cluster) is then
generated by drawing $\factorscores_{i} \sim \mathcal{N}(\scoremeans_k,\scorecovs_k)$, projecting by the factor loads, and adding variance.

We fit this generated data using the expectation-maximization algorithm
as described in the previous Section. After 58 iterations
(and a serial processing time of 40 seconds) we
reached our convergence threshold: the log probability increased by
less than $10^{-5}$ with each iteration. For this experiment we specified
the number of latent factors $J$ and the number of clustering components
$K$ to include, but below we repeat this experiment where the number
of latent factors and clustering components is not known.

We show a corner plot of the data in Figure~\ref{fig:toy-model-data}.
In this figure we have over-plotted 100000 samples generated by the
estimated factor loads $\factorloads_\textrm{est}$ and the clustering in factor 
scores $\scoremeans_\textrm{est}$ and $\scorecovs_\textrm{est}$ (weighted by $\vec\weight_\textrm{est}$)
-- without adding variance $\specificvariance_\textrm{est}$ % Ref Factor load figure 
-- in order to illustrate that the
estimated model parameters can realistically generate the data.
The estimated latent space of factor scores is shown in 
Figure~\ref{fig:toy-model-latent-space} where the colours indicate the
most likely component that the data point is associated with.

Latent factor models are by definition indeterminate in that there are
multiple solutions of the parameter vector $\vec\Psi$ which provide
an identical data-generating process. This can be trivially shown by
the fact that the factors loads and scores can be rotated with some rotation
matrix $\textbf{R}$ and generate the exact same data. As a consequence,
we cannot strictly evaluate our model fitting procedure by comparing
how closely the inferred factor loads (or scores) match the true 
variables. This is further complicated by orientation and order
of the latent factors: even if the inferred factor loads were very close
to the true factor loads, they could be ordered differently, or have
their signs reversed. Given these issues, we evaluate the search as
being successful if the log probability given the estimated model
parameters is within convergence tolerance (or better) than the
log probability given the data and the true parameters.

% Despite not matching perfectly, it generates the data? property of uniqueness?
% Factor scores?
% Some metric about clustering performance against true values?

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{experiments/toy-model-data.png}
    \caption{\todo{Caption}}
    \label{fig:toy-model-data}
\end{figure*}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{experiments/toy-model-latent-space.png}
    \caption{\todo{Caption}}
    \label{fig:toy-model-latent-space}
\end{figure*}

% Search strategy for toy model.

% Figure: toy-model-factor-loads?
% Figure: toy-model-specific-variances?

\subsection{Experiment 2: A subset of 14 chemical abundances in \Galah}
\label{sec:experiment-galah}

We use the second \Galah\ data release \citep{Buder:2018a} which
includes up to 23 chemical abundances measured for 342,682
stars. It is not practical to use all of these measurements
when performing chemical tagging. For example, our experiments
with \Galah\ data do not include lithium abundances because the
photospheric lithium varies throughout a star's lifetime. 
We first selected stars with \texttt{flag\_cannon = 0} to exclude
stars where there is reason to suspect that the stellar parameters
(e.g., $\teff$, $\logg$) are unreliable, and as a result the 
detailed chemical abundances are untrustworthy.

We included the following fourteen chemical abundances: Na, Mg, 
Si, Ca, Sc, Ti, Cr, Mn, Fe, Ni, Cu, Zn, Ba, and Eu.
By requiring that each star has \texttt{flag\_cannon = 0} and
\texttt{flag\_<x> = 0} (where \texttt{<x>} is the relevant label)
for each chemical abundance, this leaves
us with a sample of 12,798 stars each with reliable stellar
parameters ($\teff$, $\logg$) and chemical abundances.
This is relevant to the approach we use here, as the method
described here requires no missing data.
These elements are produced from multiple nucleosynthetic
pathways, and it is this subsample that we will use for 
our initial chemical experiment with \Galah\ data.

\begin{comment}
For this subset
we examined the measured abundances and
counted the number of stars with \texttt{flag\_x\_fe = 0}
for each abundance. This flag indicates that there is no known
reason to suspect that the measured chemical abundance is
unreliable.

We sorted the elemental abundances by the number of stars
with reliable measurements, from highest to lowest. Starting
with Fe, we iterated over each element and counted the number
of stars with reliable measurements of that element and all
other elements earlier in the list. We illustrate this process
in Figure~\ref{fig:galah-abundances}, where we show that
if we include the top 14 elemental abundances, we would have
12,798 stars with reliable measurements in all fourteen elements.
This is relevant to the approach we use here, as the method
described here requires no missing data.
\end{comment}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{experiments/galah-abundances.png}
    \caption{\todo{Caption}}
    \label{fig:galah-abundances}
\end{figure*}

\begin{comment}
This leaves us with a sample of 12,798 stars each with
reliable stellar parameters ($\teff$, $\logg$) and fourteen
(reliably) measured abundances: Na, Mg, Si, Ca, Sc, Ti, Cr, Mn, 
Fe, Ni, Cu, Zn, Ba, and Eu
\todo{Y} (reliable) measured abundances: \todo{Mn, Fe}.
These elements are produced from multiple nucleosynthetic
pathways, including \todo{X}. 
It is this subsample that we will use for experiments
in chemical tagging.
\end{comment}

%\subsubsection{Search strategy} \label{sec:experiment-galah-search}

The expectation-maximization procedure outlined in Section \ref{sec:methods} describes 
the process of refining our estimate of the model parameters $\vec\psi$ until
the improvement in log probability between iterations decreases below some
pre-defined threshold. This method is valid for a pre-defined number of latent
factors, and a pre-defined number of components to cluster in latent space.
In practice these values are not known. For the purposes of this work, we 
perform a grid-based search in $J$ and $K$, perform E-M for each combination of
$J$ and $K$, and record the optimal parameters.

We trial the number of latent factors from $J = 1$ to $J = 12$, and trial the
number of clustering components from $K = 1$ to $K = 40$. We re-initialise the
mixture for each new trial of $\{J, K\}$, and do not make use of results from
`nearby' mixtures. After the E-M threshold of $10^{-5}$ has been reached, we
record the log likelihood of the data given the model parameters, the
Bayesian Information Criterion \citep{bic}, 
\begin{equation}
	\textrm{BIC} = Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:bic}
\end{equation} 
\noindent{}and the pseudo-Bayesian Information Criterion \citep{pseudo-bic}
\begin{equation}
 \textrm{pseudo-BIC} = 6(1 + \gamma)\omega{}Q\log{N} - 2\log\mathcal{L}\left(\data|\vec\Psi\right) \label{eq:pseudo-bic}
\end{equation}
\noindent{}where $Q$ is the total number of model parameters in both equations,
and the unknown constants $\omega$ and $\gamma$ have constraints
$\omega \geq 1$ and $\gamma > 0$. Here we adopt $\gamma = 0.1$ and $\omega = 1$.

\todo{Discuss grid search, inferred factor loads, cluster variances, etc.}

\begin{figure}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-ll-contours.png}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-bic-contours.png}
	\includegraphics[width=0.45\textwidth]{experiments/galah-experiment-2-gridsearch-pseudobic-contours.png}
    \caption{Model performance metrics for Experiment 2 using a subset of 14 
    		 \Galah\ chemical abundances measured from 12,798 stars. The top 
		 	 panel shows the negative log-likelihood 
			 $-\log{\mathcal{L}\left(\data|\vec\Psi\right)}$ 
			 evaluated at each combination of latent factors $J$ and number 
			 of clusters $K$, the middle panel shows the Bayesian Information
			 Criterion (Equation \ref{eq:bic}), and the lower panel shows
			 the pseudo-Bayesian Information Criterion 
			 (Equation~\ref{eq:pseudo-bic}). The white marker indicates the 
			 lowest value in each panel.}
    \label{fig:experiment-2-gridsearch}
\end{figure}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{experiments/galah-experiment-1-gridsearch-factors-wrt-J-and-K.png}
	\caption{Latent factor loads inferred for models with increasing numbers
			 of latent factors and increasing numbers of cluster components
			 in Experiment~2. The latent factors have each been rotated by a
			 rotation matrix $\textbf{R}$ with entries $\{-1, 0, +1\}$ such 
			 that the estimated latent factors in each model are ordered 
			 and oriented to be as similar as the latent factors estimated from
			 the model with only one latent factor. Note that the rotation
			 matrix contains only entries $\{-1, 0, +1\}$. The latent
			 factors for each model shown here were initialised using singular value
			 decomposition on a random subset of the data.
			 }
	\label{fig:experiment-2-gridsearch-factors-init-svd}
\end{figure*}

\begin{figure*}
	\includegraphics[width=1.0\textwidth]{experiments/galah-experiment-2-gridsearch-factors-wrt-J-and-K.png}
	\caption{Latent factor loads inferred for models with increasing numbers
			 of latent factors and increasing numbers of cluster components
			 in Experiment~2. The latent factors have each been rotated by a
			 rotation matrix $\textbf{R}$ with entries $\{-1, 0, +1\}$ such 
			 that the estimated latent factors in each model are ordered 
			 and oriented to be as similar as the latent factors estimated from
			 the model with only one latent factor. Note that the rotation
			 matrix contains only entries $\{-1, 0, +1\}$. The latent
			 factors for each model shown here were initialised randomly.
			 }
	\label{fig:experiment-2-gridsearch-factors-init-svd}
\end{figure*}

\subsection{Experiment 4: A minimal set of nucleosynthetic trace elements in \Galah}

In this experiment we investigate the effects of initialising the model parameters.
Here we make use of a minimal set of nucleosynthetic trace elements
using the second \Galah\ data release \citep{Buder:2018}. Specifically, we use
reliable chemical abundance measurements of Na (an odd-Z element), Mg and Ca (alpha-elements),
Ni, Fe, and Cr (iron-peak elements), Ba and Y (s-process tracers), and Eu (an r-process tracer).
We assume that there are four latent factors ($J = 4$) and four components ($K = 4$),
and we initialise the latent factors and component parameters in different combinations.
The latent factors can be initialised in four different ways: (1) with entries drawn randomly from $\mathcal{U}\left(-1, +1\right)$, (2) with noisey draws $\mathcal{N}\left(0, 10^{-2}\right)$, (3) using singular value decomposition, and (4) using astrophysical insight to initialise the factor loads. 
\todo{When using astrophysical insight...}
The component parameters are either initialised randomly, or using K-means++ algorithm in
the latent space. For each combination of initialisation procedure, we run 30 optimisations
and record the log likelihood once the E-M procedure has reached a threshold of $10^{-8}$.

The results of this experiment are shown in Figure~\ref{fig:experiment-4-log-L}.

\section{Results} \label{sec:results}

\section{Discussion} \label{sec:discussion}

\section{Conclusions} \label{sec:conclusion}

\acknowledgements
A.~R.~C. is supported in part by Australian Research Council
Discovery Project DP160100637.
% Gaia acknowledgement?
% GALAH acknowledgement?

\software{
	\package{Astropy} \citep{astropy:v1,astropy:v2},
    \package{IPython} \citep{ipython},
    \package{matplotlib} \citep{mpl},
    \package{numpy} \citep{numpy},
    \package{scipy} \citep{scipy},
    \package{Stan} \citep{stan},
    \package{Jupyter Notebooks} \citep{jupyter-notebooks}
}

\bibliographystyle{aasjournal}
\bibliography{mcfa}

\end{document}